---
title: "PML-Project"
author: "Rekha Vaidyanathan"
date: "March 14, 2016"
output: html_document
---

<h1>Submitted by Rekha Vaidyanathan</h1>
<h2>Data Preparation</h2>

<p> Read the Training and Testing Set</p>
```{r}
trainingSet<- read.csv("pml-training.csv")
testdata<-read.csv("pml-testing.csv")
```


<p>Import useful libraries</p>
```{r}
library(randomForest)
library(caret)
library(e1071)
library(rpart)
library(rpart.plot)
library(rattle)

```
<p>Select the valid columns related to features</p>

```{r}
sensorColumns = grep(pattern = "_belt|_arm|_dumbbell|_forearm", names(trainingSet))
newTrainingSet=trainingSet[,c(sensorColumns,160)]
dim(newTrainingSet)
```

<p>Clean the Dataset by removing missing data ("NA"" data) and handle Non-Zero values </p>
```{r}
missingData=is.na(newTrainingSet)
omitcolumns=which(colSums(missingData)>12000)
T = newTrainingSet[,-omitcolumns]
dim(T)
nzv<-nearZeroVar(T,saveMetrics=TRUE)
newT<-T[,nzv$nzv==FALSE]
dim(newT)

```
<p>Treat the Test data similar to Training data.Keep only the columns obtained after treatment of data in Training set. Remove the last column containing class information</p>
```{r}
colForTest<-colnames(newT[,-53])
testingData<-testdata[colForTest]
dim(testingData)

```
<p>Coerce the data according to treated Training Set</p>
```{r}

for (i in 1:length(testingData) ) {
    for(j in 1:length(newT)) {
         if( length( grep(names(newT[i]), names(testingData)[j]) ) == 1) 
 {
             class(testingData[j]) <- class(newT[i])
         }      
     }      
 }

 dim(testingData)

```
<p>Split to training and test set with 60:40 </p>
```{r}
train<-createDataPartition(y=newT$classe,p=0.6,list=FALSE)
training<-newT[train,]
dim(training)

testing<-newT[-train,]
dim(testing)
```
<h2> Using Machine Learning for Prediction</h2>
<h3>Prediction using Random Forest Model </h3>
```{r}
fitRF<-randomForest(classe~.,data=training)
predictRF<-predict(fitRF,testing,type="class")
confMatrix<-confusionMatrix(predictRF,testing$classe)
confMatrix
```
<h3>Fit Model using CART </h3>
```{r}
modrpart<-rpart(classe~.,data=training,method="class")

```
<p>Plot Tree</p>

```{r, echo=FALSE}
 
fancyRpartPlot(modrpart)
```

<h3>Prediction using CART </h3>
```{r}
predictRpart<-predict(modrpart,testing,type="class")
conf<-confusionMatrix(predictRpart,testing$classe)
conf

```
<h3>Better Results obtained using Random Forest.
Prediction of Test Data using Random Forest Model </h3>
```{r}

predictRF<-predict(fitRF,testingData,type="class")

predictRF
```
